Comparing tool variability and numerical variability in fMRI analyses

### 2021-09-28 paper/paper.tex

\newcommand{\fslspm}{FSL-SPM\xspace}
\newcommand{\fslafni}{FSL-AFNI\xspace}
\newcommand{\afnispm}{AFNI-SPM\xspace}

@@

Computational environments are known to affect the pipeline results in neuroimaging,
presumably due to the creation, propagation, and amplification of small numerical errors.
Numerical instability is a characteristic of the pipelines which amplify numerical errors
originating in floating-point computations and then hamper reproducibility of the analyses.
% It has been widely investigated that methodological flexibility can influence the final determined areas of brain activation.
In particular, errors that highlight numerical instabilities are likely to appear as a result of changes in data analysis software packages.
This work presents a comparative assessment of the group-level fMRI analysis between tool variability and numerical variability.   
We characterize the effect of numerical perturbations on the fMRI analysis using the Fuzzy Libmath (FL) framework.
FL is a technique to instrument the mathematical functions embedded in mathematical libraries based on the Monte-Carlo Arithmetic (MCA).
We show that by taking advantage of the Fuzzy libmath framework, one can evaluate the numerical uncertainty of the pipelines. 
We (1) evaluated the numerical stability of the fMRI analysis that has been reproduced using three of the most common tools in neuroimaging, FSL, AFNI, and SPM,
and (2) compared between tool variability and numerical variability.
We found a significant numerical variability on both thresholded and unthresholded images;
results show that the order of magnitude of variations between tools was highly more than numerical variability.
Also, we found a linear correlation between Dice scores computed region-by-region on thresholded maps in both conditions.

>>>>

\newcommand{\fslspm}{FSL-SPM\xspace}
\newcommand{\fslafni}{FSL-AFNI\xspace}
\newcommand{\afnispm}{AFNI-SPM\xspace}
\newcommand{\tristan}[1]{\color{orange}\textbf{From Tristan:} #1\color{black}\xspace}

@@

Numerical and software variability broadly affect structural, diffusion, and functional MRI analyses. Numerical
variability originates in software updates or code
parallelization, whereas software variability reflects discrepancies between
models implemented in different analysis software packages. While numerical
and software variability were both shown to impact analysis outcomes, the
extent to which these sources of variability compare for a given
analysis remains understudied. This work presents a comparison of
numerical and software variability for group-level functional MRI analysis
\tristan{did we give up on individual analyses?}.
We reproduced a previous comparison between functional MRI analysis 
software packages FSL, AFNI, and SPM, which we extended to measure 
numerical variability through Monte-Carlo arithmetic. 
We find that between-tool variability is an order of magnitude higher than numerical variability.
\tristan{summarize conclusions}
